---
title: "MA308PROJ_ANOVA"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# import dataset
```{r}
library(readr)
heart <- read_csv("heart_cleveland_upload.csv")
attach(heart)
```
I was wondering the sample size of each group differs a lot, so can ANOVA test still works?


Questions:
1.parallel factor we can see the condition's difference between the label; but for continuous number, we should see the CI and mean from the condition's aspect.
2.If we use One-way ANCOVA, does the importance of oldpeak and slope equal?

#  one-way ANOVA for CP
```{r}
library(gplots)
library(HH)
library(multcomp)
library(car)
#table(cp)
aggregate(condition, by=list(cp),FUN=mean)
aggregate(condition, by=list(cp),FUN=sd)
fitcp <- aov(condition ~ cp)
summary(fitcp)
plotmeans(condition ~ cp)
#TukeyHSD(fitcp)
bartlett.test(condition ~ cp)
```


#  one-way ANOVA for slope
```{r}
#table(slope)
aggregate(condition, by=list(slope),FUN=mean)
aggregate(condition, by=list(slope),FUN=sd)
fitslope <- aov(condition ~ slope)
summary(fitslope)
plotmeans(condition ~ slope)

```

#  one-way ANOVA for thal
```{r}
#table(thal)
aggregate(condition, by=list(thal),FUN=mean)
aggregate(condition, by=list(thal),FUN=sd)
fitthal <- aov(condition ~ thal)
summary(fitthal)
plotmeans(condition ~ thal)

```

#  one-way ANOVA for ca
```{r}
#table(ca)
aggregate(condition, by=list(ca),FUN=mean)
aggregate(condition, by=list(ca),FUN=sd)
fitca <- aov(condition ~ ca)
summary(fitca)
plotmeans(condition ~ ca)

```

#  one-way ANOVA for sex
```{r}
#table(sex)
aggregate(sex, by=list(condition),FUN=mean)
aggregate(sex, by=list(condition),FUN=sd)
fitsex <- aov(sex ~ condition)
summary(fitsex)
plotmeans(sex ~ condition)

```

#  one-way ANOVA for trestbps
```{r}
#table(trestbps)
aggregate(trestbps, by=list(condition),FUN=mean)
aggregate(trestbps, by=list(condition),FUN=sd)
fittrestbps <- aov(trestbps ~ condition)
summary(fittrestbps)
plotmeans(trestbps ~ condition)

```

#  one-way ANOVA for oldpeak
```{r}
#table(oldpeak)
aggregate(oldpeak, by=list(condition),FUN=mean)
aggregate(oldpeak, by=list(condition),FUN=sd)
fitoldpeak <- aov(oldpeak ~ condition)
summary(fitoldpeak)
plotmeans(oldpeak ~ condition)
```

#  One-way ANCOVA for oldpeak and slope
```{r}
slope <- as.factor(slope)
fito_s <- aov(condition ~ oldpeak + slope)
summary(fito_s)
```

but! All anova should use the assumption as normal distribution, but the actually we can not assume...
Then let's try another one called wilcoxon signed-rank test
but actually... there occurs some questions.
(1. we count one mean and test for another; 2.not paired)
So let's turn to Mann-Whitney test
(but the sample is not large)
