---
title: "Heart Disease"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(normalizePath("."))
library(car) 
library(DT)
library(flextable)
library(tidyverse)
library(dslabs)
library(dplyr)
library(caret)
library(lubridate)
library(tidytext)
library("RColorBrewer")
library(randomForest)
library(tictoc)
library(e1071)
library(ggpubr)

```

```{r}
heart = read.csv("heart_cleveland_upload.csv")
is.null(heart)
names(heart)[ncol(heart)]="target"
datatable(heart)
heart_latent = heart
heart_latent$cp1 = as.numeric(heart$cp ==1)
heart_latent$cp2 = as.numeric(heart$cp ==2)
heart_latent$cp3 = as.numeric(heart$cp ==3)
heart_latent$restecg1 = as.numeric(heart$restecg ==1)
heart_latent$restecg2 = as.numeric(heart$restecg ==2)
heart_latent$slope1 = as.numeric(heart$slope ==1)
heart_latent$slope2 = as.numeric(heart$slope ==2)
heart_latent$ca1 = as.numeric(heart$ca ==1)
heart_latent$ca2 = as.numeric(heart$ca ==2)
heart_latent$ca3 = as.numeric(heart$ca ==3)
heart_latent$thal1 = as.numeric(heart$thal ==1)
heart_latent$thal2 = as.numeric(heart$thal ==2)
heart_latent = subset(heart_latent, select = -c(cp,restecg,slope,ca,thal))
heart_latent$target = as.factor(heart_latent$target)
levels(heart_latent$target) =make.names(c("target","target"),unique = T)

```

1.1 Data explaination:
* age: The person's age in years
* sex: The person's sex (1 = male, 0 = female)
* cp: The chest pain experienced (Value 1: typical angina, Value 2: atypical angina, Value 3: non-anginal pain, Value 4: asymptomatic)
* trestbps: The person's resting blood pressure (mm Hg on admission to the hospital)
* chol: The person's cholesterol measurement in mg/dl
* fbs: The person's fasting blood sugar (> 120 mg/dl, 1 = true; 0 = false)
* restecg: Resting electrocardiographic measurement (0 = normal, 1 = having ST-T wave abnormality, 2 = showing probable or definite left ventricular hypertrophy by Estes' criteria)
* thalach: The person's maximum heart rate achieved
* exang: Exercise induced angina (1 = yes; 0 = no)
* oldpeak: ST depression induced by exercise relative to rest ('ST' relates to positions on the ECG plot.)
* slope: the slope of the peak exercise ST segment (Value 1: upsloping, Value 2: flat, Value 3: downsloping)
* ca: The number of major vessels (0-3)
* thal: A blood disorder called thalassemia (3 = normal; 6 = fixed defect; 7 = reversable defect)
* target: Heart disease (0 = no, 1 = yes) 


1.2 Grouping the factors:
  1)Unchangable: increasing age, male gender and heredity(thalassemia).
  2)Modifiable: Smoking, high cholesterol, high blood pressure, physical inactivity, and being overweight and having diabetes
  3)Prior: high cholesterol, high blood pressure, diabetes, weight, family history and smoking



2 Model
```{r}
# Set partation: 
# train set: train_set
# Validation set: validation
set.seed("2077", sample.kind = "Rounding")
# Divide into train and validation dataset
test_index <- createDataPartition(y = heart_latent$target, times = 1, p = 0.8, list= FALSE)
temp = heart_latent[-test_index,]
validation_index = createDataPartition(y= temp$target, times = 1, p = 0.5, list = FALSE)
train_set <- heart_latent[test_index, ]
validation <- temp[validation_index, ]
test = temp[-validation_index,]

# Converting the dependent variables to factors
train_set$target <- as.factor(train_set$target)
validation$target <- as.factor(validation$target)
test$target <- as.factor(test$target)
```

2 GLM Model
    2.1 Assumption
    Since the target variable $Y_{i}$ is binary form, we assume it $ Y_{i} \sim Binomial(1,\pi(X_{i}))$
    
    2.2 Variable Selection
```{r}
fit.full = glm(target~., family = binomial(link = "logit"), data = train_set)
as_flextable(fit.full)
# Didn't Discovered Multicollinearility
vif.full = as.data.frame(vif(fit.full))
vif.full

fit.step_AIC = step(fit.full, k = 2);
fit.step_BIC = step(fit.full, k = log(nrow(heart_latent)));

as_flextable(fit.step_AIC)
as_flextable(fit.step_BIC)

fit.final = fit.step_BIC
```
From the p-values for the regression coefficients (last column), you can see that age, chol, fbs, restecg may not make a significant contribution to the equation. 


Anova table
```{r}
anova = anova(fit.full, fit.step_BIC,test = "Chisq")
flextable(data = anova)
```

So from the anova tabel it suggest the final model is better, because the p-value>0.05 so we cannot reject the null that the final model is as useful as the full model.

    2.3 Feature Importance? / Coefficient analysis
    ```{r}
coef = as.data.frame(exp(coef(fit.final)))
names(coef) = "Coefficient"
coef
```
The coeficient



```{r}
# Confidence Interval of coefficients
coef = exp(confint(fit.final))
as.data.frame(exp(confint(fit.final)))
```
    
    2.4 Model Dagonise
      2.4.1 Residuals
```{r}
# residual analysis
plot(x=predict(fit.final, type="response"),y = residuals(fit.final), xlab = "Y", ylab = "
     Deviance")
abline(h = 0)
chi.sim = rchisq(length(residuals(fit.final)), df= fit.final$df.residual)
qqplot(qchisq(ppoints(297),df = fit.final$df.residual),residuals(fit.final), xlab = "Theoretical chi-square", ylab = "Deviance")
qqline(residuals(fit.final), datax = FALSE, distribution =function(p) qchisq(p,fit.final$df.residual), probs = c(0.25, 0.75), qtype = 7, col = "red")


```
From the qq-Plot we can see the residual doesn't follow gaussian distribution. And the Shapiro-Walker test statistics for testing normality equals to 0.977 with p-value equal to 0.000089, hence we can reject the normality hypothesis at the significient levels, which coincide with the qq-plot result.

```{r}
# Influencial points
influencePlot(fit.final)
index.influ = c(91,134,178,214,257)
heart_latent[index.influ,]
```
      2.4.2 multicollinearity
```{r}
library(mctest)
#cor = as.data.frame(round(cor(train_set),digits = 3))
#(cor)
library(ggcorrplot)
cor <- cor(train_set[,-9])
ggcorrplot(cor,lab = F)

vif = as.data.frame(vif(fit.final))
vif
```

  Confussion matrix
Validation
```{r}
# confussion matrix
library("caret")
true = as.factor(validation$target)
pred = as.numeric(predict(fit.final,type = "response", validation)>0.5)
pred = as.factor(pred)
levels(pred) = levels(heart_latent$target)
confusion_matrix = confusionMatrix(data = pred, reference = true)
confusion_matrix

```
Test confussion matrix
```{r}
library("caret")
true = as.factor(test$target)
pred = as.numeric(predict(fit.final,type = "response", test)>0.5)
pred = as.factor(pred)
levels(pred) = levels(heart_latent$target)
confusion_matrix = confusionMatrix(data = pred, reference = true)
confusion_matrix

```



  logistic ROC
```{r}
library(ROCR)
library(pROC)

#### AIC

probfull <- predict(fit.final, newdata = test, type = "response")
predfull <- prediction(probfull, test$target)
AUC_full <- performance(predfull, "auc")@y.values[[1]]
plot(performance(predfull, "tpr", "fpr"), colorize = T, lwd = 3, main = "ROC Curves of GLM")

fitsex <- glm(target ~ sex, data = train_set, family = binomial())
probsex <- predict(fitsex, newdata = test, type = "response")
predsex <- prediction(probsex, test$target)
AUC_sex <- performance(predsex, "auc")@y.values[[1]]
plot(performance(predsex, "tpr", "fpr"), colorize = T, lwd = 3, main = "ROC Curves")

fittrestbps <- glm(target ~ trestbps, data = train_set, family = binomial())
probtrestbps <- predict(fittrestbps, newdata = test, type = "response")
predtrestbps <- prediction(probtrestbps, test$target)
AUC_trestbps <- performance(predtrestbps, "auc")@y.values[[1]]
plot(performance(predtrestbps, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")


fitcp3 <- glm(target ~ cp3, data = train_set, family = binomial())
probcp3 <- predict(fitcp3, newdata = test, type = "response")
predcp3 <- prediction(probcp3, test$target)
AUC_cp3 <- performance(predcp3, "auc")@y.values[[1]]
plot(performance(predcp3, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")

fitca3 <- glm(target ~ ca3, data = train_set, family = binomial())
probca3 <- predict(fitca3, newdata = test, type = "response")
predca3 <- prediction(probca3, test$target)
AUC_ca3 <- performance(predca3, "auc")@y.values[[1]]
plot(performance(predca3, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")

fitthalach <- glm(target ~ thalach, data = train_set, family = binomial())
probthalach <- predict(fitthalach, newdata = test, type = "response")
predthalach <- prediction(probthalach, test$target)
AUC_thalach <- performance(predthalach, "auc")@y.values[[1]]
plot(performance(predthalach, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")

fitthal2 <- glm(target ~ thal2, data = train_set, family = binomial())
probthal2 <- predict(fitthal2, newdata = test, type = "response")
predthal2 <- prediction(probthal2, test$target)
AUC_thal2 <- performance(predthal2, "auc")@y.values[[1]]
plot(performance(predthal2, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")

fitca1 <- glm(target ~ ca1, data = train_set, family = binomial())
probca1 <- predict(fitca1, newdata = test, type = "response")
predca1 <- prediction(probca1, test$target)
AUC_ca1 <- performance(predca1, "auc")@y.values[[1]]
plot(performance(predca1, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")

fitca2 <- glm(target ~ ca2, data = train_set, family = binomial())
probca2 <- predict(fitca2, newdata = test, type = "response")
predca2 <- prediction(probca2, test$target)
AUC_ca2 <- performance(predca2, "auc")@y.values[[1]]
plot(performance(predca2, "tpr", "fpr"), add = TRUE, colorize = T, lwd = 3, main = "ROC Curves")

cat("AUC of Logistic:",AUC_full)



AUC_ALL <- data.frame(x1="thal2", x2=AUC_thal2)
row<- c("cp3", AUC_cp3)
AUC_ALL <- rbind(AUC_ALL, row)
row<- c("thalach", AUC_thalach)
AUC_ALL <- rbind(AUC_ALL, row)
row<- c("sex", AUC_sex)
AUC_ALL <- rbind(AUC_ALL, row)
row<- c("ca2", AUC_ca2)
AUC_ALL <- rbind(AUC_ALL, row)
row<- c("trestbps", AUC_trestbps)
AUC_ALL <- rbind(AUC_ALL, row)
row<- c("ca3", AUC_ca3)
AUC_ALL <- rbind(AUC_ALL, row)
row<- c("ca1", AUC_ca1)
AUC_ALL <- rbind(AUC_ALL, row)

AUC_ALL

```
3 Mechine Learning Model
  3.1 LDA( Linear Discrimination Analysis)
```{r}

Ctrl <- trainControl(summaryFunction=twoClassSummary,
                     classProbs=TRUE)
lda_fit <- train(target ~ ., method = "lda", data = train_set, preProc = c("center","scale"),trControl=Ctrl,metric = "ROC")

lda_fit
#Best Tune
res <- lda_fit$results
res
#importance
varImp(lda_fit)

```
Diagnosis
```{r}
# LDA
lda_predict <- predict(lda_fit, validation)
lda_results <- confusionMatrix(lda_predict, validation$target)
lda_results
```

  3.2 KNN Classifer
```{r}
ctrl <- trainControl(summaryFunction=twoClassSummary,method = "cv", verboseIter = FALSE, number = 5, classProbs = T)
knnFit <- train(target ~ ., data = train_set, method = "knn", 
                preProcess = c("center","scale"),
                trControl = ctrl , tuneGrid = expand.grid(k =
                seq(1, 20, 2)), metric = "ROC")
# Best Tune
knnFit$bestTune
plot(knnFit)
res <- knnFit$results
res

# Importance
varImp(knnFit)
```
  
  3.3 Random Forest
```{r}

set.seed(2077)
control<- trainControl(summaryFunction=twoClassSummary,method = "cv", number = 5, verboseIter = FALSE, classProbs = T)
grid <-data.frame(mtry = seq(1, 10, 2))
tic(msg= " Total time for rf :: ")
rf_fit <- train(target ~ ., method = "rf", data = train_set, ntree = 20, trControl = control,
                  tuneGrid = grid, metric = "ROC")
# Best Tune
rf_fit$bestTune
plot(rf_fit)
res <- rf_fit$results
res

# Importance
varImp(rf_fit)



```

3.4 Neral Network
```{r}
library(neuralnet)
library(caret)
library(magrittr)
library(dplyr)

set.seed(3)
heart_latent$desease_yes =
  heart_latent$target =="target.1"
heart_latent$desease_no = heart_latent$target =="target"

train_set$desease_yes =
  train_set$target =="target.1"
train_set$desease_no = train_set$target =="target"

test$desease_yes =
  test$target =="target.1"
test$desease_no = test$target =="target"

network = neuralnet(desease_no+desease_yes ~ age + sex + trestbps + chol + fbs + thalach + exang + 
    oldpeak + cp1 + cp2 + cp3 + restecg1 + restecg2 + slope1 + 
    slope2 + ca1 + ca2 + ca3 + thal1 + thal2,data = train_set, hidden = 3)

plot(network)
```

Network Analysis
```{r}
#confussion matrix
net.predict = neuralnet::compute(network, test)$net.result
net.prediction = c("no desease","has desease")[apply(net.predict, 1, which.max)]
# net.prediction
predict.table = table(test$target, net.prediction, dnn = c("Actual","Predicted"))
predict.table
```

```{r}
netResult <- function(table){
  TP = table[2];FP = table[4]; FN = table[1];TN = table[3]
  sensitivity = (TP)/(TP + FN)
  accuracy = (TP+TN)/(TP + FP + FN + TN)
  precision = (TP)/(TP + FP)
  specificity = (TN)/(TN + FP)
  result = list(accuracy = accuracy,sensitivity = sensitivity, specificity = specificity)
  return(result)
}
netResult(predict.table)
```

Model selection:
1. Confussion Matrix
```{r}
# LDA
lda_predict <- predict(lda_fit, test)
lda_results <- confusionMatrix(lda_predict, test$target)
lda_results
#KNN
knnPredict <- predict(knnFit,newdata = test )
knn_results <- confusionMatrix(knnPredict, test$target )
knn_results
#Random Forest
rf_predict <- predict(rf_fit, newdata = test)
rf_results <- confusionMatrix(rf_predict, test$target)
rf_results

```


Model selection:
  2.ROC
```{r}
library(pROC)                  # plot the ROC curve
library(xgboost)      # Extreme Gradient Boosting
# lda
lda.probs = predict(lda_fit,test, type = "prob")
lda.ROC = roc(predictor=lda.probs$target, 
              response = test$target, 
              levels = rev(levels(test$target)),
              )
lda.ROC$auc
#Area under the curve: 0.8731
plot(lda.ROC,main="GBM ROC",col = "Orange")

#KNN
knn.probs = predict(knnFit,test, type = "prob")
knn.ROC = roc(predictor=knn.probs$target, 
              response = test$target, 
              levels = rev(levels(test$target)))
knn.ROC$auc
#Area under the curve: 0.8731
plot(knn.ROC,main="KNN ROC",col = "Blue",add = T)

#Randome Forest
rf.probs = predict(rf_fit,test, type = "prob")
rf.ROC = roc(predictor=rf.probs$target, 
              response = test$target, 
              levels = rev(levels(test$target)))
rf.ROC$auc
plot(rf.ROC,main="Randome Forest ROC",col = "Green", add = T)

#Area under the curve: 0.8731

legend("bottomright", lty = 1,col = c("Orange","Blue","Green"), legend = c("LDA","KNN","Random Forest"))
```

# import dataset
```{r}
library(readr)
heart <- read_csv("heart_cleveland_upload.csv")
heart_trans <- read_csv("heart_cleveland_upload.csv")
```

```{r}
library(ggcorrplot)
cor_heart <- cor(heart_trans[,1:14])
cor_heart
ggcorrplot(cor_heart,lab = F)
```

# transition
```{r}
 library(dbplyr)
 library(dplyr)
 library(plyr)
heart_trans <- heart_trans %>% 
  mutate(sex = if_else(sex == 1, "1-MALE", "0-FEMALE"),
         fbs = if_else(fbs == 1, ">120", "<=120"),
         exang = if_else(exang == 1, "YES", "NO"),
         cp = if_else(cp == 0, "0-typical angina",
                      if_else(cp == 1, "1-atypical angina", if_else(cp == 2, "2-non-anginal pain", "3-asymptomatic"))),
         restecg = if_else(restecg == 0, "0-NORMAL",
                           if_else(restecg == 1, "1-ABNORMALITY", "2-PROBABLE OR DEFINITE")),
         slope = if_else(slope == 0, "0-upsloping",
                           if_else(slope == 1, "1-flat", "2-downsloping")),
         ca = as.factor(ca),
         thal  = if_else(ca == 0, "0-normal",
                           if_else(ca == 1, "1-flxed defect", "2-reversable defect", "3")),
    
         condition = if_else(condition == 1, "YES", "NO")
         ) %>% 
  mutate_if(is.character, as.factor) %>% 
  dplyr::select(condition, sex, fbs, exang, cp, restecg, slope, ca, thal, everything())
```

# the number of the condition
```{r}
attach(heart_trans)
 library(ggplot2)
ggplot(heart_trans, aes(x=condition, fill=condition)) + 
  geom_bar() +
  xlab("Heart Disease") +
  ylab("Count")
```

# histgram age & condition
```{r}
  ggplot(heart_trans, aes(x=age,fill=condition))+
  geom_histogram()+
  scale_x_discrete() +
#    scale_fill_brewer(palette = "Blues") +
  xlab("Age") + 
  ylab("Number")+
  guides(fill = guide_legend(title = "Condition"))

  ggplot(heart_trans, aes(x=trestbps,fill=condition))+
  geom_histogram()+
  scale_x_discrete() +
#    scale_fill_brewer(palette = "Blues") +
  xlab("trestbps") + 
  ylab("Number")+
  guides(fill = guide_legend(title = "Condition"))
  
    ggplot(heart_trans, aes(x=thalach,fill=condition))+
  geom_histogram()+
  scale_x_discrete() +
#    scale_fill_brewer(palette = "Blues") +
  xlab("thalach") + 
  ylab("Number")+
  guides(fill = guide_legend(title = "Condition"))
```

```{r}
  ggplot(heart_trans, aes(x=slope,y=oldpeak,color=condition))+
  geom_point(alpha=0.7)+xlab("oldpeak") +
  ylab("slope")+
  guides(fill = guide_legend(title = ""))
```

```{r}
heart_trans %>%
  ggplot(aes(x=condition,y=chol))+
  geom_boxplot(fill="#D55E00")+
  xlab("Condition")+
  ylab("Chol")

heart_trans %>%
  ggplot(aes(x=condition,y=thalach))+
  geom_boxplot(fill="#D55E00")+
  xlab("Condition")+
  ylab("thalach")

heart_trans %>%
  ggplot(aes(x=condition,y=oldpeak))+
  geom_boxplot(fill="#D55E00")+
  xlab("Condition")+
  ylab("oldpeak")
```

```{r}
library(lmPerm)
library(multcomp)
set.seed(2077)

fit=aovp(condition~chol, data = heart, perm="Prob")
summary(fit)

fit=aovp(condition~thalach, data = heart, perm="Prob")
summary(fit)

fit=aovp(condition~oldpeak, data = heart, perm="Prob")
summary(fit)
```


```{r}
ggplot() + geom_bar(data = heart_trans,aes(x = thal, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = cp, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = ca, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = sex, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = fbs, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = restecg, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = exang, fill=factor(condition)), position="dodge")

ggplot() + geom_bar(data = heart_trans,aes(x = slope, fill=factor(condition)), position="dodge")
```

```{r}
ggplot(heart_trans, aes(x = age, y = thalach, color = condition, shape = condition))+
  geom_point()+
  geom_smooth(se = FALSE)
```


```{r}
library(gridExtra)
data1 <- heart_trans[(heart_trans$condition == "YES" & heart_trans$exang == "NO"), c("cp")]
data2 <- heart_trans[(heart_trans$condition == "YES" & heart_trans$exang == "YES"), c("cp")]
data3 <- heart_trans[(heart_trans$condition == "NO" & heart_trans$exang == "NO"), c("cp")]
data4 <- heart_trans[(heart_trans$condition == "NO" & heart_trans$exang == "YES"), c("cp")]
grid.arrange(
  ggplot() + geom_bar(data = data1,aes(x = cp), fill = c("#6495ED","#FFA500","#FF4500", "red3")) + 
    ggtitle("People with Heart Disease and exang = 0") + ylim(0,70)+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme(plot.title = element_text(hjust = 0.5,size = 10, face = "bold")),
ggplot() + geom_bar(data = data2,aes(x = cp), fill = c("#6495ED","#FFA500","#FF4500", "red3")) + ggtitle("People with Heart Disease and exang = 1") + ylim(0,70)+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ theme(plot.title = element_text(hjust = 0.5,size = 10, face = "bold")),
ncol = 2)

grid.arrange(
ggplot() + geom_bar(data = data3,aes(x = cp), fill = c("#6495ED","#FFA500","#FF4500", "red3")) + ggtitle("People without Heart Disease and exang = 0") + ylim(0,70)+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ theme(plot.title = element_text(hjust = 0.5,size = 10, face = "bold")),
ggplot() + geom_bar(data = data4,aes(x = cp) , fill = c("#6495ED","#FFA500","#FF4500", "red3")) + ggtitle("People without Heart Disease and exang = 1") + ylim(0,70)+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ theme(plot.title = element_text(hjust = 0.5,size = 10, face = "bold")),
ncol = 2)
```


Prediction
```{r}
heartPre = function(age,sex){
  health = data.frame(trestbps = rep(140,4),sex = rep(sex,4),age = rep(age,4), ca1 = 0, ca2= 0,ca3 = 0, thal2  =rep(0,4), thalach = rep(220-age,4), cp3 = 1)

  health$ca1 = c(0,1,0,0)
  health$ca2 = c(0,0,1,0)
  health$ca3 = c(0,0,0,1)
  
  predict_res = predict(fit.final,type = "response", health)
  
  cat("Hi dear, you have a chance between",min(predict_res)*100,"% and", max(predict_res)*100,"% getting a heart disease.")
}

 
heartPre(age = 65, sex = 1)
heartPre(age = 107, sex = 0)

heartPre(age = 20, sex = 1)
heartPre(age = 20, sex = 0)
```

